---
title: "Manuscript for a Data Analysis Project"
subtitle: ""
author: Katie Wells & Connor Ross
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl

---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.


```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyverse)
```

## Importing and viewing data
```{r}

## Loading the full-raw data set...
Raw_USVac <- read_csv(here("data", "raw_data" , "COVID-19_Vaccinations_in_the_United_States_Jurisdiction.csv"))


# Check data
str(Raw_USVac)

summary(Raw_USVac)

```

```{r}
# Data Wrangling

## Take out year 2023. Only want data from 2021 - 2022.

library(lubridate)

Raw_USVac$Date <- mdy(Raw_USVac$Date)


library(tidyverse)
SubData <- Raw_USVac %>%
  filter(Date > '2021-02-01' & Date < '2022-12-31')

library(ggplot2)
library(dplyr)

## Vacccine Distribution Overtime by Manufac. US Wide
DisTime <- SubData %>%
  dplyr::group_by(`Date`) %>%
  dplyr::select(`Date`, 5:8) %>%
  pivot_longer(!Date, names_to = "Vacc_Dist", values_to = "Count")

Graph_DisTime <- DisTime %>%
  drop_na(Count)%>%
  ggplot(mapping = aes(x = Date, y = Count, fill = Vacc_Dist)) +
  geom_area() +
  ggtitle("Vaccine distribution over time")
Graph_DisTime

## Vaccine Administration Overtime by Manufac. US Wide
AdminTime <- SubData %>%
  dplyr::group_by(`Date`) %>%
  dplyr::select(`Date`, 20:23) %>%
  pivot_longer(!Date, names_to = "Vacc_Admin", values_to = "Count")

Graph_AdminTime <- AdminTime %>%
  drop_na(Count)%>%
  ggplot(mapping = aes(x = Date, y = Count, fill = Vacc_Admin)) +
  geom_area() +
  ggtitle("Vaccine Administration over time")
Graph_AdminTime

SubData <- SubData %>%
  dplyr::mutate(Region = case_when(
    (`Location` %in% c("ME", "VT", "NH", "MA", "CT", "RI", "NY", "NJ", "PA", "DE", "MD")) ~ "NorthEast",
    (`Location` %in% c("DC", "WV", "VA", "KY", "NC", "TN", "AR", "LA", "MS", "AL", "GA", "FL", "SC")) ~ "SouthEast",
    (`Location` %in% c("NA", "SD", "NE", "KS", "MN", "IA", "MO", "WI", "IL", "MI", "IN", "OH")) ~ "MidWest",
    (`Location` %in% c("OK", "TX", "NM", "AZ")) ~ "SouthWest",
    (`Location` %in% c("WA", "OR", "CA", "NV", "ID", "MT", "WY", "CO", "HI", "AK")) ~ "West"
  )) %>%
  drop_na(Region)

AdminBar <- SubData %>%
  group_by(Region) %>%
  dplyr::select(`Region`, 20:23) %>%
  pivot_longer(!Region, names_to = "Vacc_Admin", values_to = "Count") %>%
  group_by(Vacc_Admin)


library(ggplot2)

Graph_AdminBar <- AdminBar %>%
  dplyr::filter(Region == "NorthEast") %>%
  dplyr::summarise(sum = Count) %>%
  ggplot(mapping = aes(x = Vacc_Admin, fill = Vacc_Admin)) +
  geom_bar()
  
Graph_AdminBar


```


# Summary/Abstract
_Write a summary of your project._


{{< pagebreak >}}


# Introduction 

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_ 

## Description of data and data source
This dataset contains US COVID-19 vaccine deliveries and administration data at national and jurisdiction level from all vaccine partners, including jurisdictional partner clinics, retail pharmacies, long-term care facilities, dialysis centers, Federal Emergency Management Agency and Health Resources and Services Administration partner sites, and federal entity facilities. It was obtained from https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-Jurisdi/unsk-b7fc. It has 109 columns and 37,500 rows of administration, distribution, and series data of COVID-19 vaccinations across the US.

## Questions/Hypotheses to be addressed
Question: What/Which COVID-19 vaccine was most popular in each region of the United States?
Hypotheses: H0 - There is not a difference in COVID-19 vaccine (across manufacturers) popularity across all the regions in the United States.
HA - There is differences in COVID-19 vaccine (manufacturer) popularity across the regions in the United States.





To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}


# Methods 

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._


## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._


## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._


## Statistical analysis
After placing each state in a region of the US, we will conduct a time series analysis to determine the change in popularity over time. Further, we intended to plot overall administration of manufacturer-specific vaccines in each region to determine popularity after adjusting for distribution and time.


{{< pagebreak >}}


# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._


@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.





## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._


@fig-result shows a scatterplot figure produced by one of the R scripts.




## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.




{{< pagebreak >}}


# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses. 

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template. 

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.


{{< pagebreak >}}

# References



