---
title: "Manuscript for a Data Analysis Project"
subtitle: ""
author: Katie Wells & Connor Ross
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl

---


```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyverse)
```

# KATIE, EVERYTHING IN THE BELOW CODE CHUNK CAN BE DELECTED (IT WAS MOVED TO 'wrangling.qmd')
## Data import and cleaning
## Importing and viewing data
```{r}

## Loading the full-raw data set...
Raw_USVac <- read_csv(here("data", "raw_data" , "COVID-19_Vaccinations_in_the_United_States_Jurisdiction.csv"))


# Check data
str(Raw_USVac)

summary(Raw_USVac)

```

# Introduction 

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_ 

## Description of data and data source
This dataset contains US COVID-19 vaccine deliveries and administration data at national and jurisdiction level from all vaccine partners, including jurisdictional partner clinics, retail pharmacies, long-term care facilities, dialysis centers, Federal Emergency Management Agency and Health Resources and Services Administration partner sites, and federal entity facilities. It was obtained from https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-Jurisdi/unsk-b7fc. It has 109 columns and 37,500 rows of administration, distribution, and series data of COVID-19 vaccinations across the US.

## Questions/Hypotheses to be addressed
Question: What/Which COVID-19 vaccine was most popular in each region of the United States?
Hypotheses: H0 - There is not a difference in COVID-19 vaccine (across manufacturers) popularity across all the regions in the United States.
HA - There is differences in COVID-19 vaccine (manufacturer) popularity across the regions in the United States.

# Summary/Abstract
_Write a summary of your project._


# KATIE, EVERYTHING IN THE BELOW CHUNK HAS BEEN MOVED AND SPLIT UP. THIS CAN LIKELY BE DELETED UNLESS YOU WANTED TO PUT THE GRAPHS/TABLES HERE. RUN ALL 1. wrangling.qmd 2. exploration.qmd & 3. fitting.qmd IN THAT ORDER TO GET THE OUTPUTS IN YOUR ENVIRONMENT TO CODE THEM TO APPEAR HERE IF THAT'S WHAT YOU WANT! REMINDER!!!!! file 'fitting.qmd' WILL TAKE A WHILE TO RUN BECAUSE WE RAN CODE TO CHECK THE MODEL WE CREATED.
```{r}
# Data Wrangling

## Take out year 2023. Only want data from 2021 - 2022.

library(lubridate)

Raw_USVac$Date <- mdy(Raw_USVac$Date)


library(tidyverse)
SubData <- Raw_USVac %>%
  filter(Date > '2021-02-01' & Date < '2022-12-31')

library(ggplot2)
library(dplyr)

## Vacccine Distribution Overtime by Manufac. US Wide
DisTime <- SubData %>%
  dplyr::group_by(`Date`) %>%
  dplyr::select(`Date`, 5:8) %>%
  pivot_longer(!Date, names_to = "Vacc_Dist", values_to = "Count")

Graph_DisTime <- DisTime %>%
  drop_na(Count)%>%
  ggplot(mapping = aes(x = Date, y = Count, fill = Vacc_Dist)) +
  geom_area() +
  ggtitle("Vaccine distribution over time")
Graph_DisTime

## Vaccine Administration Overtime by Manufac. US Wide
AdminTime <- SubData %>%
  dplyr::group_by(`Date`) %>%
  dplyr::select(`Date`, 20:23) %>%
  pivot_longer(!Date, names_to = "Vacc_Admin", values_to = "Count")

Graph_AdminTime <- AdminTime %>%
  drop_na(Count)%>%
  ggplot(mapping = aes(x = Date, y = Count, fill = Vacc_Admin)) +
  geom_area() +
  ggtitle("Vaccine Administration over time")
Graph_AdminTime

SubData <- SubData %>%
  dplyr::mutate(Region = case_when(
    (`Location` %in% c("ME", "VT", "NH", "MA", "CT", "RI", "NY", "NJ", "PA", "DE", "MD")) ~ "NorthEast",
    (`Location` %in% c("DC", "WV", "VA", "KY", "NC", "TN", "AR", "LA", "MS", "AL", "GA", "FL", "SC")) ~ "SouthEast",
    (`Location` %in% c("NA", "SD", "NE", "KS", "MN", "IA", "MO", "WI", "IL", "MI", "IN", "OH")) ~ "MidWest",
    (`Location` %in% c("OK", "TX", "NM", "AZ")) ~ "SouthWest",
    (`Location` %in% c("WA", "OR", "CA", "NV", "ID", "MT", "WY", "CO", "HI", "AK")) ~ "West"
  )) %>%
  drop_na(Region)

library(tidyverse)

AdminBar <- SubData %>%
  dplyr::select(`Region`, 20:23) %>%
  group_by(`Region`)

AdminBar3 <- SubData %>%
  dplyr::select(`Region`, 20:23) %>%
  tidyr::drop_na(everything()) %>%
  dplyr::group_by(`Region`) %>%
  dplyr::summarise(JanssenCount = sum(`Administered_Janssen`),
                   ModernaCount = sum(`Administered_Moderna`),
                   PfizerCount = sum(`Administered_Pfizer`),
                   NovavaxCount = sum(`Administered_Novavax`)) %>%
  pivot_longer(!Region, names_to = "Vacc_man", values_to = "Count")
  
  


library(ggplot2)

AdminBar3$Region <- as.factor(AdminBar3$Region)

AdminBar3$Count <- as.numeric(AdminBar3$Count)


# Southeast Region
RegionSoutheast <- AdminBar3 %>%
  dplyr::filter(Region == "SouthEast") %>%
  dplyr::select(2:3)

RegionSoutheast$Vacc_man <- as.character(RegionSoutheast$Vacc_man)

Graph_SEAdminBar <- RegionSoutheast %>%
  ggplot(mapping = aes(x = reorder(Vacc_man, -Count), y = Count, fill = Vacc_man)) +
  geom_bar(stat = 'identity') 

Graph_SEAdminBar <- Graph_SEAdminBar + geom_text(aes(label = Count)) + 
  labs(title = "Vaccines Administered in the SouthEast", x = "Vaccine Manufacters", y = "Count")
  
  
Graph_SEAdminBar




# Northeast Region
RegionNortheast <- AdminBar3 %>%
  dplyr::filter(Region == "NorthEast") %>%
  dplyr::select(2:3)

RegionNortheast$Vacc_man <- as.character(RegionNortheast$Vacc_man)

Graph_NEAdminBar <- RegionNortheast %>%
  ggplot(mapping = aes(x = reorder(Vacc_man, -Count), y = Count, fill = Vacc_man)) +
  geom_bar(stat = 'identity') 

Graph_NEAdminBar <- Graph_NEAdminBar + geom_text(aes(label = Count)) + 
  labs(title = "Vaccines Administered in the NorthEast", x = "Vaccine Manufacters", y = "Count")

Graph_NEAdminBar



# MidWest Region
RegionMidwest <- AdminBar3 %>%
  dplyr::filter(Region == "MidWest") %>%
  dplyr::select(2:3)

RegionMidwest$Vacc_man <- as.character(RegionMidwest$Vacc_man)

Graph_MWAdminBar <- RegionMidwest %>%
  ggplot(mapping = aes(x = reorder(Vacc_man, -Count), y = Count, fill = Vacc_man)) +
  geom_bar(stat = 'identity') 

Graph_MWAdminBar <- Graph_MWAdminBar + geom_text(aes(label = Count)) + 
  labs(title = "Vaccines Administered in the Midwest", x = "Vaccine Manufacters", y = "Count")

Graph_MWAdminBar


# SouthWest Region
RegionSouthwest <- AdminBar3 %>%
  dplyr::filter(Region == "SouthWest") %>%
  dplyr::select(2:3)

RegionSouthwest$Vacc_man <- as.character(RegionSouthwest$Vacc_man)

Graph_SWAdminBar <- RegionSouthwest %>%
  ggplot(mapping = aes(x = reorder(Vacc_man, -Count), y = Count, fill = Vacc_man)) +
  geom_bar(stat = 'identity') 

Graph_SWAdminBar <- Graph_SWAdminBar + geom_text(aes(label = Count)) + 
  labs(title = "Vaccines Administered in the Southwest", x = "Vaccine Manufacters", y = "Count")

Graph_SWAdminBar

# West Region
RegionWest <- AdminBar3 %>%
  dplyr::filter(Region == "West") %>%
  dplyr::select(2:3)

RegionWest$Vacc_man <- as.character(RegionWest$Vacc_man)

Graph_WAdminBar <- RegionWest %>%
  ggplot(mapping = aes(x = reorder(Vacc_man, -Count), y = Count, fill = Vacc_man)) +
  geom_bar(stat = 'identity') 

Graph_WAdminBar <- Graph_WAdminBar + geom_text(aes(label = Count)) + 
  labs(title = "Vaccines Administered in the West", x = "Vaccine Manufacters", y = "Count")

Graph_WAdminBar

```
We plan to break this up later into separate sections/R scripts in a cleaner version but this will work for our preliminary purposes. Here we show the frequency of the distribution and administration of the different vaccine manufacturers over 2021-2022. We also show the number of each type of vaccine administered in each region that we designated. In out statistical analysis we plan to bring in population data for each region and discover trends in vaccine administration, distribution, and preference in relation to population.



{{< pagebreak >}}



# Methods 

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._


## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._



## Statistical analysis
After placing each state in a region of the US, we will conduct a time series analysis to determine the change in popularity over time. Further, we intended to plot overall administration of manufacturer-specific vaccines in each region to determine popularity after adjusting for distribution and time.


{{< pagebreak >}}


# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._



## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._




## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._






{{< pagebreak >}}


# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_




{{< pagebreak >}}

# References



